Near and mid-infrared properties of known $z\geq5$ Quasars

### 2018-08-27 tex/older_tex/WW4C_VHzQ_20180729.tex
README.md and tex directory added

\title[Very high-$z$ Quasars]
        {Near and Mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
       {Nicholas P. Ross$^{1}$\thanks{Corresponding Author: npross@roe.ac.uk} and Nicholas J. G. Cross$^{1}$
\\ 
$^1$Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\
}

\maketitle
\begin{abstract}
In this paper, we, for the first time since the discovery of $z\geq5$
quasars, assemble all spectroscopically confirmed very high redshift
quasars in one catalogue.  In particular we present the near
($zZyYJHK_{s}$ and $K$) infrared and mid-infrared (WISE) properties of
all 424 spectroscopically confirmed redshift $z\geq5.00$ quasars.
Using archival public WFCAM/UKIRT and VIRCAM/VISTA data we check for
photometric variability in the near-infrared that might be expected
from Super-Eddington accretion and find {\it blah}.  We present a
comprehensive series of colour-redshift and colour-colour plots and
make inferences into the hot dust properties of the very high-redshift
quasar population. Extrapolating the known quasar luminosity function
we suggest that $x$\% of the possibly detected $z\geq5$ quasars in the
current datasets have been discovered.
\end{abstract}

@@

\iffalse
\section*{TO DOs}
{\bf For NPR: }
\begin{itemize}
\item Concentrate on reporting NIR and W1/W2 figures and tables
\item SED dust plots 
\item Filter curve plot with LSST, Euclid and even wide JWST filters...; 
\item Check for NIR and MIR variabilty with the Wu quasar
\item Check with SpIES/SHELA (but only 3.4/4.6$\mu$m; 
%%\item Almost certainly want to compile M$_{\rm BH}$; THIS CAN WAIT, otherwise is massive paper-creep
%%\item May well wanna try and get M$_{\star}$ too...!! (yuck); THIS CAN WAIT, otherwise is massive paper-creep
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Very high redshift quasars (VH$z$Q; defined here to have redshifts
$z\geq5.00$) are excellent probes of the early Universe. This includes
studies of the Epoch of Reionization for hydrogen \citep[see e.g.][for
reviews]{Fan2006review, Mortlock2016}, the formation and build-up of
supermassive black holes \citep[e.g., ][]{Rees1984, WyitheLoeb2003,
Volonteri2010, Agarwal2016, Valiante2018, Latif2018} and early metal
enrichment \citep[see e.g., ][]{Simcoe2012, Chen2017, Bosman2017}.

Super-Eddington (also called super-critical) accretion is a viable
physical mechanism to explain the high luminosity and rapid growth of
supermassive black holes in the early universe
\citep[e.g.,][]{AlexanderNatarajan2014, MadauHaardtDotti2014,
Volonteri2015, Pezzulli2016, Lupi2016, Pezzulli2017, Takeo2018}. Thus,
one might well expect VH$z$Qs to vary in luminosity as they
potentially go through phases of Super-Eddington accretion and these
signatures of photometric variablity should be looked for, especially
when the e.g. near-infrared $K$-band is sampling rest-frame
$\approx$3670, 3145, 2750\AA\ at redshifts $z=5, 6$ and 7,
respectively.

Quasars are also known to be prodigious emitters of infrared emission,
thought to be from the thermal emission of dust grains heated by
continuum emission from the accretion disc
\citep[e.g.,][]{Richards2006b, Leipski2014, Hill2014, Hickox2017}.
Observations in the mid-infrared, e.g. $\sim$3-30$\mu$m allow
discrimination between AGN\footnote{Historically, ``quasars'' and
``Active Galactic Nuclei (AGN)'' have described different
luminosity/classes of objects, but here we use these terms
interchangeably (with a preference for quasar) in recognition of the
fact that they both describe accreting supermassive black holes
\citep[e.g.][]{Haardt2016book}.}  and passive galaxies due to the
1.6$\mu$m ``bump'' entering the MIR at $z\approx0.8-0.9$ \citep[e.g.,
][]{Wright1994, Sawicki2002, Lacy2004, Stern2005, Richards2006b,
Timlin2016} as well as AGN and star-forming galaxies due to the
presence of Polycyclic Aromatic Hydrocarbon (PAHs) at $\lambda >3\mu$m
\citep[e.g., ][]{Yan2007, Tielens2008}.

@@

This paper can be considered an update of \citet{Blain2013} and also
an extension of parts of \citet{Banados2016}, with the latter study
reporting WISE W1, W2, W3 and W4 magnitudes for Panoramic Survey
Telescope and Rapid Response System 1 \citep[Pan-STARRS1,
PS1;][]{Kaiser2002, Kaiser2010}, but with no further investigation
into the reddest WISE waveband for the VH$z$Qs.  \citet{Banados2016}
reports and investigates the W1, W2 and W3 properties of quasars at $z
> 5.6$. We chose redshift $z=5.00$ as our lower redshift limit due to
a combination of garnishing a large sample, adequately spanning
physical properties (e.g. luminosity, age of the Universe) and to
incorporate what knowledge we have gained over the last couple of
decades since $z>5$ quasars were discovered.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%%trim=l b r t
%  \includegraphics[width=18.6cm, clip,trim=14mm 4mm 10mm 10mm]
  \includegraphics[width=8.6cm, clip,trim=32mm 4mm 32mm 10mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/SEDs/filters_vs_QSOstars_20180704.pdf}
  \centering
  \vspace{-12pt}
  \caption[]
  {The spectral bands used by different survey telescopes and that are relevant here.
    The $grizy$ filters are from the Pan-STARRS survey. The $JHK$ are from 
    UKIRT/WFCAM, while $K_{s}$ is a VISTA/VIRCAM filter. The 
    The narrow W-band centered at $\lambda\approx14,500$\AA\ is a CFHT/Wircam filter. 
    The WISE passbands  W1-4 are also presented.
    The quasar spectrum is a composite based on \citet{VdB2001} and 
    \citet{Banados2016}. The L and T dwarf spectra are from \citet{Cushing2006}. 
  }
  \label{fig:filters}
\end{figure}

\vspace{-16pt}
\section{Data}
In Table~\ref{tab:THE_TABLE} we present our dataset and 
we have assembled this in the following manner.  First
we compile the list of all known, spectroscopically confirmed quasars
from the literature. Most of these objects are easily identified by
their broad Ly$\alpha$ emission line, \nv emission and characteristic
shape blueward of 1215\AA\ observed. As we shall see, some of the more
recently discovered objects are close to the galaxy luminosity
function characteristic luminosity $M^{*}$, and some have relatively
weak or maybe even completely absorbed Ly$\alpha$ \citep[e.g. Figures
7 and 10 in][]{Banados2016}. We leave aside detailed investigation and
discussion into espectral features and line strengths, and take as given 
the published spectra.

We then obtain optical, near-infrared and mid-infrared photometry for the
spectral dataset. The optical data comes from the 
Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 
survey \citep{Chambers2016} and the Dark Energy Camera \citep{Flaugher2015}, 
including the Dark Energy Survey \citep[DES;][]{Flaugher2005} and 
the Dark Energy Camera Legacy Survey \citep[DECaLS]{Dey2018}.

Recently, Hyper Suprime-Cam \citep[HSC; ][]{Miyazaki2018} has embarked
on the Subaru Strategic Program \citep[SSP; ][]{Aihara2018a}
delivering exquisite multi-band photometry data
\citep[][]{Aihara2018b} the five broad bands $grizy$.

The near-infrared data comes from two primary sources. First, the Wide
Field Camera \citep[WFCAM; ][]{Casali2007} on the the United Kingdom
Infra-Red Telescope (UKIRT) primarly, but not exclusively as part of
the UKIRT Infrared Deep Sky Survey \citep[UKIDSS; ][]{Lawrence2007}.
And second, data from the VIRCAM (VISTA InfraRed CAMera) on the VISTA
\citep[Visible and Infrared Survey Telescope for
Astronomy;][]{Emerson2006, Dalton2006}.

We also obtain mid-infrared, $\lambda=3-30\mu$m wavelength data from the
the Wide-Field Infrared Survey Explorer \citep[WISE;][]{Wright2010, Cutri2013} mission.
WISE mapped the sky in 4 passbands, in bands centered at wavelengths of 3.4, 4.6, 12, and 23$\mu$m. 
In total the release all sky ``ALLWISE'' catalog, contains nearly 750 million detections at high-significance\footnote{
\href{http://wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}{http://wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}}.

Figure~\ref{fig:filters} displays the wavelength and normalised transmission 
of the filters in question.

@@

\subsection{Spectroscopy} 
We have obtained a list of 424 spectroscopically confirmed 
quasars with redshifts $z\geq5.00$.

In Table~\ref{tab:THE_TABLE}  we give the discovery reference for the
VH$z$Qs noting that some objects were discovered independently and
contemporaneously.  The redshifts for the VH$z$Qs generally come from
the measurement of broad UV/optical emission lines. Where 
there are far infra-red emission lines e.g. \cii~158 micron, we report 
these, but at the level of our current analysis broadline redshifts are
sufficient.

Specifically, we use data from: \citet{Fan2000}, \citet{Fan2001c},
\citet{Fan2003}, \citet{Fan2004}, \citet{Mahabal2005},
\citet{Cool2006}, \citet{Fan2006}, \citet{Goto2006},
\citet{McGreer2006}, \citet{Carilli2007}, \citet{Kurk2007},
\citet{Stern2007}, \citet{Venemans2007}, \citet{Willott2007},
\citet{Jiang2008}, \citet{Wang2008}, \citet{Jiang2009}, \citet{Kurk2009}, 
\citet{Mortlock2009}, \citet{Willott2009}, \citet{Carilli2010}, 
\citet{Wang2010}, \citet{Willott2010a}, \citet{Willott2010b}, 
\citet{DeRosa2011}, \citet{Mortlock2011}, \citet{Wang2011}, 
\citet{Zeimann2011}, \citet{Morganson2012}, \citet{Venemans2012}, 
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Wang2013}, 
\citet{Willott2013b}, \citet{Banados2014}, \citet{Calura2014}, 
\citet{Leipski2014}, \citet{Banados2015a}, \citet{Banados2015b}, 
\citet{Becker2015}, \citet{Carnall2015}, \citet{Jiang2015}, 
\citet{Kashikawa2015}, \citet{Kim2015}, \citet{Reed2015}, 
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Willott2015}, 
\citet{Wu2015}, \citet{Venemans2016}, \citet{Wang2016_WISE}, 
\citet{Matsuoka2016}, \citet{WangR2016}, \citet{Mortlock2011},
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Venemans2013},
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Banados2016},
\citet{Matsuoka2016}, \citet{Reed2017}, \citet{Wang2017},
\citet{Mazzucchelli2017}, \citet{Ikeda2017}, \citet{Tang2017},
\citet{Koptelova2017}, \citet{Banados2018}, \citet{Matsuoka2018a} 
and \citet{Matsuoka2018b}.

Table~\ref{tab:THE_TABLE} gives the salient details for the objects
used in this study. We use all the $z\geq5.00$ quasars that
have been discovered and spectroscopically confirmed as of the time of
writing (2018 June). We report near-infrared ($yYJHK$-bands)
and mid-infrared (WISE W1/2/3/4) photometry and give our caculcated 
$M_{1450}$.

\begin{table}
\begin{tabular}{l r l}
\hline  \hline
Survey   & \# VH$z$Qs & Notes/Survey reference  \\
\hline  
  ATLAS      &     4    &  \citet{Shanks2015} \\
  CFHQS     &   20    &  \citet{Willott2007} \\
  DELS        &     2    &  \citet{Dey2018} \\
  ELAIS       &     1    &  \citet{Vaisanen2000} \\
  FIRST       &     1    &  \citet{Becker1995} \\
%  HSC         &     8    &  \\
  IMS          &     1     &  \citet{Kim2015} \\
  MMT        &   12    &  \citet{McGreer2013} \\
  NDWFS    &     1     &  \citet{JD1999} \\
  PSO          &   84    &   \citet{Kaiser2002, Kaiser2010} \\
  RD           &     1       &  \cite{Mahabal2005} \\
  SDSS        & 156       & \citet{EDR} \\
  SDUV       &   20       & \citet{Yang2017} \\
  SDWISE    &   27    &     \citet{Wang2016} \\
  SHELLQs  &  $^{a}$63   &  \citet{Matsuoka2016}\\  %% includes 8 objects from the first HSC paper..
  ULAS       &   10   & \citet{Lawrence2007} \\
  VDES       &   11   &  \citet{Reed2017} \\
  VIK          &     9   &  \citet{Edge2013} \\
  VIMOS     &     1   &  \citet{LeFevre2003} \\
\hline  \hline
\end{tabular}
\caption{The number of VH$z$Q from given surveys, with the key survey or telescope reference. 
$^{a}$Includes 8 objects with a HyperSuprimeCam \citep[HSC; ][]{Miyazaki2018}  
designation.
}
      \label{tab:surveys}
\end{table}

@@

    We query and return the mean PSF magnitudes from the $grizy$ filters
    ({\tt MeanPSFMag}) which are in the AB system for our 424 VH$z$Q
    sample. Details of our SQL and links to the main tables are given
    in Appendix~\ref{sec:PS1_SQL}.

    \iffalse
    \subsubsection{DECam} 
    The Dark Energy Camera \citep[DECam][]{Flaugher2015} is 
    is a wide-field imager with a 2.2 degree diameter field of view 
    mounted at the prime focus of the Victor M. Blanco 4 m telescope 
    at the Cerro Tololo International Observatory. We use data 
    from both the Dark Energy Survey \citep[DES; ][]{DES2016} and 
    the DESI Legacy Imaging Survey \citep[DECaLS; ][]{Dey2018}. 
    \fi

\subsection{Near-infrared photometry}~\label{sec:NIR_data} 
Due to their selection of being very faint/undetected in the
observed-frame optical but bright in the observed frame near-infrared,
VH$z$Qs are generally detected in the near-infrared $yYJHK$-bands
\citep[$\approx$0.98-2.38$\mu$m; e.g., ][]{Peth2011}.

@@

@@

The data was processed using a forced photometry, or matched-aperture
photometry. A pipeline for this has been developed by WFAU, but
has not yet been incorporated into the main VDFS pipeline. Full details of the
matched-aperture pipeline (MAP) will appear in a forthcoming paper, Cross et al.
2018, in prep, and has also been discussed in \citet{Cross2013}.  
%% and \citet{Cross2017} -- Can't find Cross et al. 2017 in NASA/ADS.
We will give the pertinent points of the MAP and specific details of these data sets below.

    \subsubsection{UKIDSS LAS} 
    In a bit more detail, use:\\
    dr10plus\\
    lasSource table (merged catalogue), nearest object only; radius 2.0'' \\
    424 results returned. Of which...

    \subsubsection{VISTA VHS} 
    In a bit more detail, use:\\
    VH5DR5 \\
    sourceTable (merged catalogue), nearest object only; radius 2.0'' \\
    424 results returned.  Of which...

@@

    \subsubsection{Setup Matched Aperture Product}

@@

    \subsubsection{Processing MAP data and Forced photometry} 
    A set of apertureIDs were created in \verb+MapApertureIDshighzQsoMap+, based on
    the \verb+RequiredMatchedApertureProduct+ {\bf selection} value. This may
    seem unnecessary layer since it is a one-to-one match with \verb+finalQsoCatalogue+,
    but the layer is a general abstraction that allows multiple sources of
    catalogue, and multiple types\ldots
    Entries added to \verb+MapSurveyTables+ $\Rightarrow$ not pertinent

    CASU's \verb+imcore_list+ was run on each frame assigned to the programme with
    the input list equal to objects selected to be within \ldots of the multiframe
    centre. In this case, where there are 424 QSOs spread across the sky, there is
    typically 1 object per multiframe, but many have two and one,
    multiframeID=2614475 in the VSA, has 3 objects. The output of \verb+imcore_list+
    is a multi-extension FITS binary table, with identical columns to the CASU FITS
    binary tables produced using the CASU \verb+imcore+, the basis of all WFCAM and
    VISTA survey products and also familiar to many astronomers through the INTWFS, 
    VST-ATLAS and VPHAS.     {\it this needs tidied up, citations added...}
    Since forced photometry can be run
    in many ways on a particular image, a single science frame is not
    necessarily linked to a single MAP catalogue, so the \verb+MapFrameStatus+ table
    links images ({\bf multiframeID}) to catalogues ({\bf catalogueID}) and gives
    information on what processing stage they are at ({\bf bitProcessingFlag}),
    whether they have been ingested ({\bf isIngested}), or what quality control has
    been run on them ({\bf ppErrBitsStatus}).

@@

@@

%%
We can gain a good appreciation for what these missions will discover
by collating the datasets we currently have.

@@

\begin{table*}

@@

   \caption{Adapted from Table 9 of \citet{Peth2011}. 

>>>>

\title[Very high-$z$ Quasars]
        {Near and mid-infrared properties of known $z\geq5$ Quasars}
\author[Ross \& Cross]
       {Nicholas P. Ross$^{1}$\thanks{Corresponding Author: npross@roe.ac.uk} and Nicholas J. G. Cross$^{1}$
\\ 
$^1$Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh, EH9 3HJ, United Kingdom\\
}

\maketitle
\begin{abstract}
Very high redshift quasars are excellent probes of physics the early
Universe.  Understanding how billion solar mass black holes
formed, evolve and accrete in the first billion years is a key
outstanding problem in contemporary astrophysics, and one that wil lbe
directly addressed by upcoming missions such as ESA {\it Euclid} and
the {\it James Webb Space Telescope}.
%%
We can gain a good appreciation for what these missions will discover
by collating the datasets we currently have, and in this paper, we,
for the first time since the discovery of $z\geq5.00$ quasars assemble all spectroscopically confirmed very high redshift quasars in
one catalogue.
%%
In particular we present the near ($z/Z, y/Y, J, H, K_{s}$ and $K$)
infrared and mid-infrared (WISE) properties of all 424
spectroscopically known redshift $z\geq5$ quasars.  Using archival public
WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric variability
in the near-infrared and find blah.  We present a comprehensive series
of colour-redshift and colour-colour plots and make inferences into
the hot dust properties of the very high-redshift quasar
population. Extrapolating the known quasar luminosity function we
suggest that 80\% of the possibly detected $z\geq5$ quasars in the
current datasets have been discovered.
\end{abstract}

@@

\iffalse
\section*{TO DOs}
{\bf For NPR: }
\begin{itemize}
\item Almost certainly want to compile M$_{\rm BH}$; hmmmm....
\item May well wanna try and get M$_{\star}$ too...!! (yuck); 
\item Filter curve plot with LSST, Euclid and even wide JWST filters...; 
\item Check with SpIES/SHELA (but only 3.4/4.6$\mu$m; 
\item Concentrate on NIR and W1/W2
\item SED dust plot with just W1/W2...
\item Check vs. Assef R90/C75 catalog.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%  S E C T I O  N   1         S E C T I O  N   1           S E C T I O  N   1       S E C T I O  N   1
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Very high redshift quasars (VH$z$Q; and defined here to have redshifts
$z\geq5.00$) are excellent probes of the early Universe. This includes
studies of the Epoch of Reionization for hydrogen \citep[see e.g.][for
reviews]{Fan2006review, Mortlock2016}, the formation and build-up of
supermassive black holes \citep[e.g., ][]{Rees1984, WyitheLoeb2003,
Volonteri2010, Agarwal2016, Valiante2018, Latif2018} and early metal
enrichment \citep[see e.g., ][]{Simcoe2012, Chen2017, Bosman2017}.

Super-Eddington (also called super-critical) accretion is a viable
physical mechanism to explain the high luminosity and rapid growth of
supermassive black holes in the early universe
\citep[e.g.,][]{AlexanderNatarajan2014, MadauHaardtDotti2014,
Volonteri2015, Pezzulli2016, Lupi2016} Thus, one might well expect
VH$z$Qs to vary in luminosity as they potentially go through phases of
Super-Eddington accretion and these should be looked for, especially
when the e.g.  near-infrared $K$-band is sampling rest-frame
$\approx$3670, 3145, 2750\AA\ at redshifts $z=5, 6$ and 7,
respectively.

The build-up of SMBHs via accretion is important due to the desire to
understand the evolution of the black hole mass function (BHMF) and
since there is a remarkable similarity between cosmic star-formation
rate density (SFRD) and black hole accretion rate (BHAR). The SFRD and
BHAR trace each other \citep[with a normalisation factor of
$\sim3000$,][]{Willott2013b, MadauDickinson2014} across redshifts
$z=0-4$.  However, new results \citep[e.g., ][]{Vito2018a, Calhau2018}
suggest that the SFRD and BHAR do {\it not} have the same evolutionary form
at higher redshift, $z>4$. The physical reasons for this are currently very poorly
understood.  For the high-$z$ QSOs that we can measure BHAR and SFR, most
have BHAR $\gg 10^{-3}$ SFR, indeed {\it Herschel}-detected sources at
$z\simeq.8$ have $L_{\rm SF} \simeq L_{\rm AGN}$ \citep[e.g.,
][]{Netzer2014} and {\it Herschel}-detected quasars show no evolution
in the $M_{\rm BH} - M_{\star}$ relation from $z\sim2$ to $z\sim0$.
where $M_{\rm BH}, M_{\rm bulge}, M_{\star}$ are the SMBH mass,
stellar mass of the bulge and galaxy total stellar mass,
respectively. The evolution and potential importance of the SFRD and BHAR 
link is a key outstanding problem in contemporary astrophysics, 
potentially explaining the galaxy-black hole scaling relations. 
However, there may also simply be a non-causal origin of the black-hole 
galaxy scaling relations \citep{Jahnke2011}.

Quasars are also known to be prodigious emitters of infrared emission,
thought to be from the thermal emission of dust grains heated by
continuum emission from the accretion disc \citep[][]{Hill2014,
Hickox2017}.  Observations in the mid-infrared, e.g. $\sim$3-30$\mu$m
allow discrimination between AGN\footnote{Historically, ``quasars'' and
``Active Galactic Nuclei (AGN)'' have described different
luminosity/classes of objects, but here we use these terms
interchangeably (with a preference for quasar) in recognition of the
fact that they both describe accreting supermassive black holes
\citep[e.g.][]{Haardt2016book}.}  and passive galaxies due to the
1.6$\mu$m ``bump'' entering the MIR at $z\approx0.8-0.9$ \citep[e.g.,
][]{Wright1994, Sawicki2002, Lacy2004, Stern2005, Richards2006b,
Timlin2016} as well as AGN and star-forming galaxies due to the
presence of Polycyclic Aromatic Hydrocarbon (PAHs) at $\lambda >3\mu$m
\citep[e.g., ][]{Yan2007, Tielens2008}.

@@

We chose redshift $z=5.00$ as our lower redshift limit due to a combination 
of garnishing a large sample, adequately spanning physical properties 
(e.g. luminosity, age of the Universe) and to incorporate what knowledge 
we have gained over the last couple of decades since $z>5$ quasars were 
discovered.

This paper can be considered an update of \citet{Blain2013} and also
an extension of parts of \citet{Banados2016}, with the latter study reporting
WISE W1, W2, W3 and W4 magnitudes for Panoramic Survey Telescope and Rapid Response System 1 
\citep[Pan-STARRS1, PS1;][]{Kaiser2002, Kaiser2010}, but with no  
further investigation into the reddest WISE waveband for the VH$z$Qs.
%%
\citet{Banados2016} reports and investigates the W1, W2 and W3 properties
of quasars at $z > 5.6$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%  S E C T I O  N   2         S E C T I O  N   2           S E C T I O  N   2       S E C T I O  N   2
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data}

\begin{figure*}
%%trim=l b r t
  \includegraphics[width=18.6cm, clip,trim=14mm 4mm 10mm 10mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/SEDs/filters_vs_QSOstars_20180704.pdf}
  \centering
  \caption[]
  {The spectral bands used by different survey telescopes and that are relevant here.}
  \label{fig:filters}
\end{figure*}

We assemble our dataset in a relatively straightforwad manner.  First
we compile the list of all known, spectroscopically confirmed quasars
from the literature. Most of these objects are easily identified by
their broad Ly$\alpha$ emission line, \nv emission and characteristic
shape blueward of 1215\AA\ observed. As we shall see, some of the more
recently discovered objects are close to the galaxy luminosity
function characteristic luminosity $M^{*}$, and some have relatively
weak or maybe even completely absorbed Ly$\alpha$ \citep[e.g. Figures
7 and 10 in][]{Banados2016}. We leave aside detailed investigation and
discussion into espectral features and line strengths, and take as given 
the published spectra.

We then obtain optical, near-infrared and mid-infrared photometry for the
spectral dataset.

The optical data comes from the 
Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 
survey \citep{Chambers2016} and the Dark Energy Camera \citep{Flaugher2015}, 
including the Dark Energy Survey \citep[DES;][]{Flaugher2005} and 
the Dark Energy Camera Legacy Survey \citep[DECaLS]{Dey2018}.

The near-infrared data comes from the Wide Field Camera (WFCAM) on 
the United Kingdom Infra-Red Telescope

\subsection{Very high redshift quasars}
In Table~\ref{tab:THE_TABLE} we give the discovery reference for the
VH$z$Qs noting that some objects were discovered independently and
contemporaneously.  The redshifts for the VH$z$Qs generally come from
the measurement of broad UV/optical emission lines. Where 
there are far infra-red emission lines e.g. \cii~158 micron, we report 
these, but at the level of our current analysis broadline redshifts are
sufficient.

Specifically, we use data from: \citet{Fan2000}, \citet{Fan2001c},
\citet{Fan2003} , \citet{Fan2004}, \citet{Mahabal2005},
\citet{Cool2006}, \citet{Fan2006}, \citet{Goto2006},
\citet{McGreer2006}, \citet{Carilli2007}, \citet{Kurk2007},
\citet{Stern2007}, \citet{Venemans2007}, \citet{Willott2007},
\citet{Jiang2008}, \citet{Wang2008}, \citet{Jiang2009}, \citet{Kurk2009}, 
\citet{Mortlock2009}, \citet{Willott2009}, \citet{Carilli2010}, 
\citet{Wang2010}, \citet{Willott2010a}, \citet{Willott2010b}, 
\citet{DeRosa2011}, \citet{Mortlock2011}, \citet{Wang2011}, 
\citet{Zeimann2011}, \citet{Morganson2012}, \citet{Venemans2012}, 
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Wang2013}, 
\citet{Willott2013b}, \citet{Banados2014}, \citet{Calura2014}, 
\citet{Leipski2014}, \citet{Banados2015a}, \citet{Banados2015b}, 
\citet{Becker2015}, \citet{Carnall2015}, \citet{Jiang2015}, 
\citet{Kashikawa2015}, \citet{Kim2015}, \citet{Reed2015}, 
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Willott2015}, 
\citet{Wu2015}, \citet{Venemans2016}, \citet{Wang2016_WISE}, 
\citet{Matsuoka2016}, \citet{WangR2016}, \citet{Mortlock2011},
\citet{McGreer2013}, \citet{Venemans2013}, \citet{Venemans2013},
\citet{Venemans2015a}, \citet{Venemans2015b}, \citet{Banados2016},
\citet{Matsuoka2016}, \citet{Reed2017}, \citet{Wang2017},
\citet{Mazzucchelli2017}, \citet{Ikeda2017}, \citet{Tang2017},
\citet{Koptelova2017}, \citet{Banados2018}, \citet{Matsuoka2018a} 
and \citet{Matsuoka2018b}.

@@

Table~\ref{tab:THE_TABLE} gives the salient details for the objects
used in this study. We use all the $z\geq5.00$ quasars that
have been discovered and spectroscopically confirmed as of the time of
writing (2018 June). We report near-infrared ($yYJHK$-bands)
and mid-infrared (WISE W1/2/3/4) photometry and give our caculcated 
$M_{1450}$

\begin{table}
\begin{tabular}{l r l}
\hline  \hline
Survey   & \# VH$z$Qs & Notes  \\
\hline  
  ATLAS      &    4    &  \\
  CFHQS     &  20    &  \\
  DELS        &    2    &  \\
  ELAIS       &    1    &  \\
  FIRST       &    1    &  \\
  HSC         &    8    &  \\
  IMS          &   1     &  \\
  MMT        &  12    &  \\
  NDWFS    &  1      &  \\
  PSO          &    84  & \\
  RD           &  1       & \\
  SDSS        & 156       & \\
  SOUV       & 20       & \\
  SDWISE    & 27& \\
  SHELLQs  & 55 & \\
  ULAS       & 10& \\
  VDES       & 11& \\
  VIK          & 9 &\\
  VIMOS     & 1& \\
\hline  \hline
\end{tabular}
\caption{ATLAS \citet{Shanks2015}; }
      \label{tab:surveys}
\end{table}

@@

    We query and return the mean PSF magnitudes from the $grizy$ filters
    ({\tt MeanPSFMag}) which are in the AB system for our 424 VH$z$Q
    sample. Details of our SQL and links to the main tables are given
    in~\ref{sec:PS1_SQL}.

    \subsubsection{DECam}

%\newpage
\subsection{Near-infrared photometry}
Due to their selection of being very faint/undetected in the
observed-frame optical but bright in the observed frame near-infrared,
VH$z$Qs are generally detected in the near-infrared $yYJHK$-bands
\citep[$\approx$0.98-2.38$\mu$m; e.g., ][]{Peth2011}.

@@

\section{NIR Data}

@@

The data was processed using a forced photometry, or matched-aperture
photometry. A pipeline for this has been developed by WFAU, but
has not yet been incorporated into the main VDFS pipeline. Full details of the
matched-aperture pipeline (MAP) will appear in a forthcoming paper, Cross et al.
2018, in prep, and has also been discussed in \citet{Cross2013} and \citet{Cross2017}.
We will give the pertinent points of the MAP and specific details of these data sets below.

\subsection{Selection of NIR}

@@

\subsection{Setup Matched Aperture Product}

@@

\subsection{Processing MAP data}

A set of apertureIDs were created in \verb+MapApertureIDshighzQsoMap+, based on
the \verb+RequiredMatchedApertureProduct+ {\bf selection} value. This may
seem unnecessary layer since it is a one-to-one match with \verb+finalQsoCatalogue+,
but the layer is a general abstraction that allows multiple sources of
catalogue, and multiple types\ldots
Entries added to \verb+MapSurveyTables+ $\Rightarrow$ not pertinent

\subsubsection{Forced photometry}
CASU's \verb+imcore_list+ was run on each frame assigned to the programme with
the input list equal to objects selected to be within \ldots of the multiframe
centre. In this case, where there are 424 QSOs spread across the sky, there is
typically 1 object per multiframe, but many have two and one,
multiframeID=2614475 in the VSA, has 3 objects. The output of \verb+imcore_list+
is a multi-extension FITS binary table, with identical columns to the CASU FITS
binary tables produced using the CASU \verb+imcore+, the basis of all WFCAM and
VISTA survey products and also familiar to many astronomers through the INTWFS
\citep{}, and VST-ATLAS and -VPHAS \citep{}. Since forced photometry can be run
in many ways on a particular image, a single science frame is not
necessarily linked to a single MAP catalogue, so the \verb+MapFrameStatus+ table
links images ({\bf multiframeID}) to catalogues ({\bf catalogueID}) and gives
information on what processing stage they are at ({\bf bitProcessingFlag}),
whether they have been ingested ({\bf isIngested}), or what quality control has
been run on them ({\bf ppErrBitsStatus}).

@@

\subsubsection{\ldots}

@@

@@

\begin{table*}[h*]

@@

   \caption{Adapted from Table 9 of \citet{Peth11}. 
